---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# helperbench

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

helperbench evaluates whether language models can consistently perform a simple refactoring task. Models are situated in a working directory with tools to explore a codebase and asked to extract repeated code into a helper function. The eval measures whether models make the requested change without breaking unit tests.

helperbench is implemented in [ellmer](https://ellmer.tidyverse.org/) with [vitals](https://vitals.tidyverse.org/).

## Installation

helperbench is implemented as an R package for ease of installation:

```r
pak::pak("simonpcouch/helperbench")
```

Load it with:

```{r}
library(helperbench)
```

helperbench will not go to CRAN.

## Example

```{r}
#| echo: false
library(tibble)
```

The package ships with a data frame `helper_results` that contains evaluation results:

```{r}
library(dplyr)

sample_n(helper_results, 10)
```

See `helper_task()` for more on how to run this eval yourself.

## How to add a new model benchmark

All models are evaluated via [OpenRouter](https://openrouter.ai/) for
consistency. Adding a new model benchmark involves four steps:

### 1. Register the model

Add a new row to `clients_to_evaluate()` in `R/run-eval.R` with the
OpenRouter model ID and a snake_case name:

```r
clients_to_evaluate <- function() {
  tibble::tribble(
    ~client, ~name,
    # ... existing models ...
    'ellmer::chat_openrouter(model = "provider/model-name")', "model_name"
  )
}
```

### 2. Generate the eval script

Run `write_all_eval_files()` to regenerate all scripts (including your new
one) in `inst/runs/scripts/`:

```r
devtools::load_all()
write_all_eval_files()
```

This creates `inst/runs/scripts/model_name.R` with the boilerplate to run the
evaluation and save the task result.

### 3. Run the evaluation

Source the generated script:

```r
source("inst/runs/scripts/model_name.R")
```

This evaluates the model across all samples and epochs, saving the result to
`inst/runs/tasks/tsk_model_name.rda`.

### 4. Update the results dataset

In `data-raw/helper_results.R`, add entries for the new model in each of the
`case_when()` blocks:

- **Display name**: Map the snake_case name to a human-readable label
  (e.g., `model_name == "model_name" ~ "Model Name"`)
- **Type**: Categorize as `"Frontier"`, `"Budget"`, or `"Local"`
- **Cost**: Add the per-million-token input and output costs from OpenRouter

Then rebuild the dataset:

```r
source("data-raw/helper_results.R")
```
