---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# helperbench

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

helperbench evaluates whether language models can consistently perform a simple refactoring task. Models are situated in a working directory with tools to explore a codebase and asked to extract repeated code into a helper function. The eval measures whether models make the requested change without breaking unit tests.

helperbench is implemented in [ellmer](https://ellmer.tidyverse.org/) with [vitals](https://vitals.tidyverse.org/).

## Installation

helperbench is implemented as an R package for ease of installation:

```r
pak::pak("simonpcouch/helperbench")
```

Load it with:

```{r}
library(helperbench)
```

helperbench will not go to CRAN.

## Example

```{r}
#| echo: false
library(tibble)
```

The package ships with a data frame `helper_results` that contains evaluation results:

```{r}
library(dplyr)

sample_n(helper_results, 10)
```

See `helper_task()` for more on how to run this eval yourself.
